A new method for genome-wide marker development and typing holds great promise for primate population genetics

### 2012-12-12 trunk/paper/paper.tex
Fixed introduction

\title{RADseq Works in Primates, Dammit.}
\author{
        Christina M. Bergey
			\and
			Luca Pozzi
			\and
			Todd R. Disotell
			\and
			Andrew S. Burrell
	}
\date{\today}

@@

\textcolor{red}{[What is mol. primatology? Explain importance of developing markers across taxa in order to study primates at different levels (population, phylogeography, phylogenetics)]} During the first decades of molecular primatology, an impediment to researchers was the need to develop and type polymorphic markers in a taxon of interest. The markers that resulted from this time-consuming and expensive task were often uninformative when applied outside of the population used in their design, necessitating further rounds of primer design or microsatellite assays, for example (Davey et al., 2011). Due to the bottleneck caused by the inefficiency of marker discovery, many population genetic or phylogenetic studies in molecular primatology have been based on one or few loci, usually mitochondrial DNA or few microsatellites (Ting and Sterner, 2012). Such inferences can reliably give the evolutionary history of those particular regions of the genome, but they fail to adequately capture the complete complex history of the population given the mosaic nature of genomic evolution (Degnan and Rosenberg, 2009; Maddison, 1997; Edwards, 2009; Maddison and Knowles, 2006). Adequate resolution depends on high marker density, and until recently that goal has been out of reach for many primate researchers (Edwards, 2009).

@@

In the present study, we summarize the RAD-seq method in brief and note the many and varied applications since its development. We go on to discuss reasons that make RAD-seq promising for primatologists, and then demonstrate the technique in 5 primates: a lemur, New World monkey, Old World monkey, and two apes.

@@

The following is a summary of the RAD tag library preparation of Etter et al., 2011 (\textcolor{red}{Fig. 1}). The RAD-seq library preparation begins when genomic DNA from multiple individuals is digested with a restriction enzyme, such as \emph{EcoRI} or \emph{PspXI} (1a). The P1 adapter is then ligated to the fragments, connected to the sticky end at the restriction enzyme cut site. The P1 adapter contains an amplification site for PCR, an Illumina sequencing priming site, and an individual-specific barcode of five basepairs (1b). Once the barcode has been added, fragments from multiple individuals can be pooled, and the DNA is randomly sheared with a sonicator to have a length distribution under 1 kilobase (1c). To select for reads that are suitable for sequencing on the Illumina platform, the sheared samples are size selected via agarose gel electrophoresis, extracting fragments between 300 and 500 bp in length. The second adapter, P2, is a Y adapter meaning its two halves are complementary for only part of their length (1d). It is ligated to the fragments and then the fragments are amplified via PCR (1e). Because the second adapter has divergent ends, the reverse amplification primer is unable to bind until after the forward amplification primer has filled in its complementary sequence. This ensures that only RAD tags ligated to P1 are able to amplify. After few cycles of PCR to minimize the risk of introducing PCR artifacts or biases, the library is ready for final clean-up, quality control, and sequencing.

>>>>

\title{A new method for genome-wide marker development and typing holds great promise for primate population genetics}
\author{
        Christina M. Bergey
			\and
			Luca Pozzi
			\and
			Todd R. Disotell
			\and
			Andrew S. Burrell
	}
\date{\today}

@@

In molecular primatology, as in all fields of molecular biology, the development of variable genetic markers is essential for the study of organisms at different levels, from population genetic to phylogeographic to phylogenetic research. During the first decades of the discipline, an impediment to researchers was the need to develop and type polymorphic markers in a taxon of interest. The markers that resulted from this time-consuming and expensive task were often uninformative when applied outside of the population used in their design, necessitating further rounds of primer design or microsatellite assays, for example (Davey et al., 2011). Due to the bottleneck caused by the inefficiency of marker discovery, many population genetic or phylogenetic studies in molecular primatology have been based on one or few loci, usually mitochondrial DNA or a few microsatellites (Ting and Sterner, 2012). Such inferences can reliably give the evolutionary history of those particular regions of the genome, but they fail to adequately capture the complete complex history of the population given the mosaic nature of genomic evolution (Degnan and Rosenberg, 2009; Maddison, 1997; Edwards, 2009; Maddison and Knowles, 2006). Adequate resolution depends on high marker density, and until recently that goal has been out of reach for many primate researchers (Edwards, 2009).

@@

In the present study, we summarize the RAD-seq method in brief and note the many and varied applications since its development. We demonstrate the technique in 5 primates: a lemur, New World monkey, Old World monkey, and two apes, and discuss the features that make RAD-seq promising for primatologists.

@@

The following is a summary of the RAD tag library preparation protocol of Etter et al., 2011 (\textcolor{red}{Fig. 1}). The RAD-seq library preparation begins when genomic DNA is digested with a restriction enzyme, such as \emph{EcoRI} or \emph{PspXI} (1a). The P1 adapter is then ligated to the fragments, connected to the sticky end at the restriction enzyme cut site. The P1 adapter contains an amplification site for PCR, an Illumina sequencing priming site, and an individual-specific barcode of five basepairs (1b). Once the barcode has been added, fragments from multiple individuals can be pooled, and the DNA is randomly sheared with a sonicator to have a length distribution under 1 kilobase (1c). To select for reads that are suitable for sequencing on the Illumina platform, the sheared samples are size selected via agarose gel electrophoresis, extracting fragments between 300 and 500 bp in length. The second adapter, P2, is a Y adapter meaning its two halves are complementary for only part of their length (1d). It is ligated to the fragments and then the fragments are amplified via PCR (1e). Because the second adapter has divergent ends, the reverse amplification primer is unable to bind until after the forward amplification primer has filled in its complementary sequence. This ensures that only RAD tags ligated to P1 are able to amplify. After few cycles of PCR to minimize the risk of introducing PCR artifacts or biases, the library is ready for final clean-up, quality control, and sequencing.
